### 2024-09-18 paper/paper.qmd
please work

---
title: "How did Covid-19 affect the number of marriages"
subtitle: "Evidence from the Toronto"
author: 
  - Rohan Alexander
thanks: "Code and data are available at: https://github.com/RohanAlexander/marriage."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @kjlaskjasdnaf.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-asdfafdesadfsdfas), from @palmerpenguins.

```{r}
#| label: fig-asdfafdesadfsdfas
#| fig-cap: Number of marriage licenses in Toronto, by civic center (2011 - 2024)
#| echo: false

data |> 
  ggplot(aes(x = date, 
             y = marriage_licenses,
             color = civic_centre)) +
  geom_point() + 
  geom_vline(xintercept = lubridate::ymd("2020-03-13"),
              linetype = "dashed") +
  theme_minimal() +
  labs(x = "Date",
       y = "Number of marriage licenses",
       color = "Location")
```

Talk more about it.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

\newpage

# References